---
title: "R script for visualization and statistics analysis 'Time Series of Coastal Oceanographic Data in the Southeastern Indian Ocean, Indonesia'"
output: html_notebook
---
# Import the data and needed package in R.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
library(writexl)

# import data
getwd()
df <- read_xlsx("df.xlsx", sheet = "1")

nrow(df)
df <- na.omit(df)
unique(df$Depth)   #factor: 0m, 10m, 20m
df$Date <- as.Date(df$Date)
df$Depth <- as.factor(df$Depth)
df$Temp <- as.numeric(df$Temp)
df$Sal <- as.numeric(df$Sal)
df$DO <- as.numeric(df$DO)
df$pH <- as.numeric(df$pH)
df$Transparency <- as.numeric(df$Transparency)
```


# Descriptve statistics result (table 2)
We attempt to summarize and describe the features of a dataset, providing a snapshot of each environmental characteristics (Temperature, salinity, chlorophyll-a, pH, DO, and water column transparency).
 
```{r}
df$Depth <- as.factor(df$Depth)

# Custom function to calculate the mode
mode_function <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Using dplyr for summarization
summary_data <- df %>%
  group_by(Depth) %>%
  summarise(
    mean_temp = mean(Temp, na.rm = TRUE),
    median_temp = median(Temp, na.rm = TRUE),
    mode_temp = mode_function(Temp),
    min_temp = min(Temp, na.rm = TRUE),
    max_temp = max(Temp, na.rm = TRUE),
    sd_temp = sd(Temp, na.rm = TRUE),
    n_Temp = n(), 
    mean_salinity = mean(Sal, na.rm = TRUE),
    median_salinity = median(Sal, na.rm = TRUE),
    mode_salinity = mode_function(Sal),
    sd_salinity = sd(Sal, na.rm = TRUE),
    min_sal = min(Sal, na.rm = TRUE),
    max_sal = max(Sal, na.rm = TRUE),
    mean_Chl_a = mean(`Chl-a`, na.rm = TRUE),
    median_Chl_a = median(`Chl-a`, na.rm = TRUE),
    mode_Chl_a = mode_function(`Chl-a`),
    sd_Chl_a = sd(`Chl-a`, na.rm = TRUE),
    min_Chl_a = min(`Chl-a`, na.rm = TRUE),
    max_Chl_a = max(`Chl-a`, na.rm = TRUE),
    mean_pH = mean(pH, na.rm = TRUE),
    median_pH = median(pH, na.rm = TRUE),
    mode_pH = mode_function(pH),
    sd_pH = sd(pH, na.rm = TRUE),
    min_pH = min(pH, na.rm = TRUE),
    max_pH = max(pH, na.rm = TRUE),
    mean_DO = mean(DO, na.rm = TRUE),
    median_DO = median(DO, na.rm = TRUE),
    mode_DO = mode_function(DO),
    sd_DO = sd(DO, na.rm = TRUE),
    min_DO = min(DO, na.rm = TRUE),
    max_DO = max(DO, na.rm = TRUE),
    mean_transparency = mean(Transparency, na.rm = TRUE),
    median_transparency = median(Transparency, na.rm = TRUE),
    mode_transparency = mode_function(Transparency),
    sd_transparency = sd(Transparency, na.rm = TRUE),
    min_transparency = min(Transparency, na.rm = TRUE),
    max_transparency = max(Transparency, na.rm = TRUE),
    n_transparency = n()
  )
print(summary_data)

```

# Visualization each environmental variable
We aim to visualize a 2-year time series of data collection (between September 2021 and August 2024).

```{r}
# Temperature
ggplot(df, aes(x = format(Date, "%Y-%m"), y = Temp , fill = Depth)) + 
  geom_boxplot() +
  facet_grid(Depth ~ ., scales = "free_y") +
  labs(title = "Temperature",
       x = "Date (yyyy-mm)",
       y = "Temperature",
       fill = "Depth (m)") +
  scale_fill_manual(values = c("lightblue", "lightgreen", "lightpink")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), 
    axis.text.y = element_text(size = 12),                         
    axis.title.x = element_text(size = 14, margin = margin(t = 10)),  # Adds margin to x-axis title
    axis.title.y = element_text(size = 14, margin = margin(r = 10)),  # Adds margin to y-axis title
    plot.title = element_text(hjust = 0.5, size = 16),                # Centers and enlarges title
    plot.title.position = "plot",                                     # Ensures title is centered on the plot
    legend.position = "bottom",                                       
    legend.direction = "horizontal",
    strip.text.y = element_text(angle = 0, size = 12)  
        )

```

```{r}
# Salinity
ggplot(df, aes(x = format(Date, "%Y-%m"), y = Sal , fill = Depth)) + 
  geom_boxplot() +
  facet_grid(Depth ~ ., scales = "free_y") +
  labs(title = "Salinity",
       x = "Date (yyyy-mm)",
       y = "Salinity",
       fill = "Depth (m)") +
  scale_fill_manual(values = c("lightblue", "lightgreen", "lightpink")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), 
    axis.text.y = element_text(size = 12),                         
    axis.title.x = element_text(size = 14, margin = margin(t = 10)),  # Adds margin to x-axis title
    axis.title.y = element_text(size = 14, margin = margin(r = 10)),  # Adds margin to y-axis title
    plot.title = element_text(hjust = 0.5, size = 16),                # Centers and enlarges title
    plot.title.position = "plot",                                     # Ensures title is centered on the plot
    legend.position = "bottom",                                       
    legend.direction = "horizontal",
    strip.text.y = element_text(angle = 0, size = 12)  
        )

```

```{r}
# Chlorophyll-a

ggplot(df, aes(x = format(Date, "%Y-%m"), y = `Chl-a` , fill = Depth)) + 
  geom_boxplot() +
  facet_grid(Depth ~ ., scales = "free_y") +
  labs(title = "Chlorophyll-a",
       x = "Date (yyyy-mm)",
       y = "Chlorophyll-a",
       fill = "Depth (m)") +
  scale_fill_manual(values = c("lightblue", "lightgreen", "lightpink")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), 
    axis.text.y = element_text(size = 12),                         
    axis.title.x = element_text(size = 14, margin = margin(t = 10)),  # Adds margin to x-axis title
    axis.title.y = element_text(size = 14, margin = margin(r = 10)),  # Adds margin to y-axis title
    plot.title = element_text(hjust = 0.5, size = 16),                # Centers and enlarges title
    plot.title.position = "plot",                                     # Ensures title is centered on the plot
    legend.position = "bottom",                                       
    legend.direction = "horizontal",
    strip.text.y = element_text(angle = 0, size = 12)  
        )

```

```{r}
# pH

ggplot(df, aes(x = format(Date, "%Y-%m"), y = pH , fill = Depth)) + 
  geom_boxplot() +
  facet_grid(Depth ~ ., scales = "free_y") +
  labs(title = "pH",
       x = "Date (yyyy-mm)",
       y = "pH",
       fill = "Depth (m)") +
  scale_fill_manual(values = c("lightblue", "lightgreen", "lightpink")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), 
    axis.text.y = element_text(size = 12),                         
    axis.title.x = element_text(size = 14, margin = margin(t = 10)),  # Adds margin to x-axis title
    axis.title.y = element_text(size = 14, margin = margin(r = 10)),  # Adds margin to y-axis title
    plot.title = element_text(hjust = 0.5, size = 16),                # Centers and enlarges title
    plot.title.position = "plot",                                     # Ensures title is centered on the plot
    legend.position = "bottom",                                       
    legend.direction = "horizontal",
    strip.text.y = element_text(angle = 0, size = 12)  
        )
```

```{r}
# DO

ggplot(df, aes(x = format(Date, "%Y-%m"), y = DO , fill = Depth)) + 
  geom_boxplot() +
  facet_grid(Depth ~ ., scales = "free_y") +
  labs(title = "Disolved oxygen",
       x = "Date (yyyy-mm)",
       y = "DO",
       fill = "Depth (m)") +
  scale_fill_manual(values = c("lightblue", "lightgreen", "lightpink")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), 
    axis.text.y = element_text(size = 12),                         
    axis.title.x = element_text(size = 14, margin = margin(t = 10)),  # Adds margin to x-axis title
    axis.title.y = element_text(size = 14, margin = margin(r = 10)),  # Adds margin to y-axis title
    plot.title = element_text(hjust = 0.5, size = 16),                # Centers and enlarges title
    plot.title.position = "plot",                                     # Ensures title is centered on the plot
    legend.position = "bottom",                                       
    legend.direction = "horizontal",
    strip.text.y = element_text(angle = 0, size = 12)  
        )
```

```{r}
# water column transparency

df1 <- filter(df, df$Depth == 0)  #only use data in water depth = 0m
str(df1)
unique(df1$Depth) 

ggplot(df1, aes(x = format(Date, "%Y-%m"), y = df1$Transparency)) +
  geom_boxplot() +
  labs(title = "Transparency",
       x = "Date",
       y = "Transparency") +
  scale_fill_manual(values = c("lightblue")) +
  scale_y_discrete(breaks = seq(1, max(df1$Transparency, na.rm = TRUE), by = 0.5)) + # Set y-axis breaks
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        plot.title = element_text(hjust = 0.5))

```

# Technical validation 

```{r}
# Temperature

his_temp <-  ggplot(df, aes(x = Temp)) +
  geom_histogram(binwidth = 1, fill = "light blue", color = "black") +
  labs(title = "",
       x = "Temperature (°C)",
       y = "Frequency") +
  theme_minimal()

his_temp
```
```{r}
# Salinity
his_sal <-  ggplot(df, aes(x = Sal)) +
  geom_histogram(binwidth = 1, fill = "light blue", color = "black") +
  labs(title = "",
       x = "Salinity (PSU)",
       y = "Frequency") +
  theme_minimal()

his_sal
```

```{r}
# Chlorophyll-a

his_Clo <-  ggplot(df, aes(x = log(`Chl-a`))) +
  geom_histogram(binwidth = 0.5, fill = "light blue", color = "black") +
  labs(title = "",
       x = "Chlorophyll-a (mg/L)",
       y = "Frequency") +
  theme_minimal()

his_Clo
```

```{r}
# pH

his_pH <-  ggplot(df, aes(x = log(pH))) +
  geom_histogram(binwidth = 0.015, fill = "light blue", color = "black") +
  labs(title = "",
       x = "pH",
       y = "Frequency") +
  theme_minimal()

his_pH
```


```{r}
# DO

his_DO <-  ggplot(df, aes(x = log(DO))) +
  geom_histogram(binwidth = 0.025, fill = "light blue", color = "black") +
  labs(title = "",
       x = "DO",
       y = "Frequency") +
  theme_minimal()

his_DO
```

```{r}
# Water column transparency

his_trans <-  ggplot(df, aes(x = log(Transparency))) +
  geom_histogram(binwidth = 0.35, fill = "light blue", color = "black") +
  labs(title = "",
       x = "Transparency",
       y = "Frequency") +
  theme_minimal()

his_trans
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
